### 2015-07-02 narrative/intro.tex
Added bio and minor restructuring of introductionZ

@@

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogeneous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.
%
{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.  Our approach in this proposal is to enable researchers
to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.

@@

The overarching idea is that data from applications can be characterized by different
levels of importance and different level of importance over time. 
Codes such as the XGC1 simulation from C. S. Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system. Due to resource limitations, the simulation will only write about 10 PB, but
when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.

@@

One of the key aspects of our approach is to allow the users to define the
utility of their data; in other-words, the system should provide a
mechanism such that, for each data object, a user can specify pieces of the object
which have more utility (more importance).
Clearly there are many research challenges faced with
this approach including: 1) How do we make this simple enough so that users can
do this, 2) Can we define a model to allow the system to make
autonomic decisions to save the data which has the highest utility, and 3) Can
we allow the back-end storage system to interact with this model so that
it can optimize not just a single application but across all of the
applications.
 %
  In our simple scenario, a user wants to write 1 PB of data every hour
for checkpoint/restart, and they wish to write 500TB of data every 15
minutes to save analysis and visualization results which have already been reduced
using in situ reduction techniques. At this stage, they will ask the system to
write their data, and they would like to get an estimated time which they can
then figure out, through a series of rules, whether they want to write out all
of the data, postpone the write until a later time, or write a reduced amount
of data, along with some code
container to allow them to regenerate the data with a specified level of
accuracy.  Furthermore, they realize that the analysis data will be saved on
either the parallel file system, long-term tape storage, or on the campaign storage,
which is a medium-term storage
area with lower bandwidth than the parallel file system.
For each piece of data, the user will specify the lifetime so
that the data with the highest utility is kept around for the longest
period of time, but the data which has the least utility might be migrated to
lower layers of the storage hierarchy more quickly.

@@

\subsection{Storage System Perspective}
\label{subsec:storage-perspective}
From the perspective of the storage system, it is necessary to manage the data from
a given user request amongst all of the requests from all users, and try to optimize the
entire system accordingly. Today the problem is that when a user writes there is a
tremendous amount of interference from other users on the system who may be
writing or reading at the same time. By requiring a certain quality
of service the system must potentially lock out users with lower currency
they want to place in the reading or writing of their data. The storage
system must also automatically move data from one tier of the storage to
another, without affecting the quality of service that was offered at request
time. The storage system must be able to organize
data amongst all of the tiers of storage in a known, consistent way.

One of the clear research objectives of this project is to understand how the
storage system can interact with the middleware (application-aware) piece so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. Clearly we need to maintain 
QoS such that fuzzy estimates of the time it will take to write/read data from
the different layers informs any migration/eviction decisions, and clearly it
is critical that the storage system can manage the data across all of the
layers. We will use learning techniques to automatically migrate data
sub-objects across the hierarchy as well as to eventually evict data which is
large and has very low utility.

>>>>

{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.

@@

Our approach targets the expected characteristics of exascale storage hardware.
The storage layer will be partitioned into multiple heterogeneous tiers with
vastly different performance characteristics. The layers will be further
differentiated by the constraints on capacity and data lifetime within each
layer. Tape archival storage will still maintain data long term, but access to
this data will be orders of magnitude slower than the next layer and naively
accessing data from archival storage will greatly impact productivity.
%
Our approach in this proposal is to enable researchers to place more knowledge
into the SSIO system so that data can be recast from a bunch of bytes to
information which the system can then place and optimize the time to gain
knowledge from exascale simulations.  We first motivate this by a problem our
team has been working on in the past six months, and then recast this problem
into a series of research questions and challenges which must be addressed in
order to optimize the SSIO system for exascale science.

@@

Take for example the latest use case for the XGC1 simulation from C. S. Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at ALCF, NERSC, and OLCF). They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system. An expert team was assembled from the user group, the middleware I/O team to help with the application I/O (using ADIOS), the local experts of the Lustre file-system and the tape archival system (HPSS). Due to physical resource limitations, the simulation was restricted to write about 10~PB initially, but then considering the time to write and read this much data on the tape archive system, and the financial costs to store it, this plan was cut again considerably. Note that such time and cost information was not known by the user; the HPSS expert had to inform everyone else about it. The user team had to make hard decisions of which data to save from the originally planned 100~PB. The file system experts told about the then-current file system limitations in size, performance and purge policy. The user then decided to write 0.5~PB per day (the individual daily simulation runs were and are going to be separated by many days or weeks), 5~PB in total, storing maximum of 2~PB of data on disks at a time and storing the rest on tape.

This forced the team to decide which pieces of data to save from the large amount of data; in 
other words, they had to prioritize which were the most important pieces and save only those. They had to further break down the still large data into two categories, data which they thought they would touch after one or two days of runs are completed and after the whole simulation is going to be completed (after ten days of runs), and what they would do with these different categories of data (the access pattern of the large data).

They then had to develop new application specific data reduction techniques, and encode that into the application code itself on top of ADIOS, the middleware I/O layer. They then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces were on tape, and also which pieces were moved over to PPPL for immediate analysis by the user.

This solution worked well for the first day of run but after the second day more problems occurred: the variability of the time it took to write the data made the I/O prohibitively costly, because even though the project had reserved the majority of resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. There was no method for the application or the user to monitor and predict this resource usage and to dynamically change to write less data when the I/O time was going to be longer. The greatest problem, however, comes after the simulation is going to be completed: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other
timesteps will be on the file system. There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.

The overarching idea of this proposal is that data from applications can be characterized by different
levels of importance and different levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset. Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.

@@

One of the key aspects of our approach is to allow the users to define the
utility of their data; in other-words, the system should provide a
mechanism such that, for each data object, a user can specify pieces of the object
which have more utility (more importance).
There are many research challenges faced with
this approach including: 1) How do we make this simple enough so that users can
do this, 2) Can we define a model to allow the system to make
autonomic decisions to save the data which has the highest utility, and 3) Can
we allow the back-end storage system to interact with this model so that
it can optimize not just a single application but across all of the
applications.
 %
  In our simple scenario, a user wants to write 1~PB of data every hour
for checkpoint/restart, and they wish to write 500~TB of data every 15
minutes to save data for analysis and visualization. At this stage, they will ask the system to
write the data, and they would like to get an estimated time from which they can
then figure out, through a series of rules, whether they want to write out all
of the data, postpone the write until a later time, or write a reduced amount
of data using in situ reduction techniques, along with some code
container that allows for regenerating the data with a specified level of
accuracy.  Furthermore, they realize that the analysis data will be saved on
either the parallel file system, long-term tape storage, or on the campaign storage,
which is a medium-term storage
area with lower bandwidth than the parallel file system.
For each piece of data, the user will specify the lifetime so
that the data with the highest utility is kept around in fast storage for the longest
period of time, while the data with the least utility may be the first to be migrated to
lower layers of the storage hierarchy.

@@

\subsection{Storage System Perspective}
\label{subsec:storage-perspective}
From the perspective of the storage system, it is necessary to manage the data from
a given user request amongst all of the requests from all users, and try to optimize the
entire system accordingly. Today the problem is that when a user writes there is a
tremendous amount of interference from other users on the system who may be
writing or reading at the same time. By requiring a certain quality
of service the {\bf {\color {red} system must potentially lock out users with lower currency
they want to place in the reading or writing of their data}}. The storage
system must also automatically move data from one tier of the storage to
another, without affecting the quality of service that was offered at request
time. The storage system must be able to organize
data amongst all of the tiers of storage in a known, consistent way.

One of the main research objectives of this project is to understand how the
storage system can interact with the middleware (which is application-aware)  so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. We need to maintain 
QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
the different layers to inform any migration/eviction decisions, an it
is also critical that the storage system can manage (discover and retrieve) the data across all of the
layers automatically, without user intervention. We will use learning techniques to automatically migrate data
sub-objects across the hierarchy as well as to eventually evict data which is
large and has very low utility.
